|

### 2025-12-03 manuscript/pre_print.tex
Preprint ready for submission

\usepackage[font=scriptsize]{caption}
\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Resting energy expenditure of women with and without polycystic ovary syndrome: a systematic review and meta-analysis},
  pdfauthor={Richard Kirwan; Leigh Peele; Greg Nuckols; Georgia Kohlhoff; Hannah Cabré; Alyssa Olenick; James Steele},
  pdfkeywords={Polycystic ovary syndrome, PCOS, resting energy
expenditure, REE, basal metabolic rate, BMR},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Resting energy expenditure of women with and without polycystic
ovary syndrome: a systematic review and meta-analysis\thanks{Preprint,
please cite as: Kirwan, R., Peele, L., Nuckols, G., Kohlhoff, G., Cabré,
H., Olenick, A., and Steele, J. (2025). Resting energy expenditure of
women with and without polycystic ovary syndrome: a systematic review
and meta-analysis. medrχiv DOI: TO ADD. Address for correspondence:
james@steele-research.com}}

\author[1]{Richard Kirwan}
\author[2]{Leigh Peele}
\author[2]{Greg Nuckols}
\author[3]{Georgia Kohlhoff}
\author[4]{Hannah Cabré}
\author[5]{Alyssa Olenick}
\author[6,2,7]{James Steele}

\affil[1]{Research Institute for Sports and Exercise Sciences, Liverpool
John Moores Universtiy, Liverpool, UK}
\affil[2]{MacroFactor, Stronger by Science LLC, Raleigh, North Carolina,
USA}
\affil[3]{Flourishing Health, Liverpool, UK}
\affil[4]{Reproductive Endocrinology and Women's Health Research
Program, Pennington Biomedical Research Center, Baton Rouge, Louisiana,
USA}
\affil[5]{Dr Alyssa Olenick LLC}
\affil[6]{Steele Research Limited, Eastleigh, Hampshire, UK}
\affil[7]{School of Health and Biomedical Sciences, Royal Melbourne
Institute of Technology, Melbourne, VIC, Australia}

\date{December 1, 2025}
\begin{document}
\maketitle
\begin{abstract}
Context: Polycystic ovary syndrome (PCOS) is common in reproductive-age
women, who often have higher BMI classification. This is assumed to stem
from lower resting energy expenditure (REE), influencing lifestyle
intervention guidelines. However, evidence for reduced REE in women with
PCOS compared with those without is inconsistent. Objective: To
systematically search and meta-analyse the existing literature to
estimate and describe the difference in REE between women with and
without PCOS. Data Sources: A systematic search was conducted using
PubMed, Medline and Web of Science databases of published research from
January 1990 to January 2025. Study Selection: Studies that measured REE
in women living with PCOS, both with and without control arms of women
without PCOS, were included. Data Extraction: Bibliometric, demographic,
and REE data was extracted by one investigator and checked in
triplicate. Data Synthesis: Thirteen studies were included in a Bayesian
arm-based multiple condition comparison (i.e., network) type
meta-analysis model with informative priors to compare both mean REE,
and between person variation in REE, between women with and without
PCOS. Mean REE differed between groups by 30 kcal/day {[}95\% quantile
interval: -47 to 113 kcal/day{]} and the contrast ratio for between
person standard deviations was 0.98 {[}95\% quantile interval: 0.71 to
1.33{]}. Conclusions: These findings indicate that REE does not
meaningfully differ between women with and without PCOS. Group-level
differences in resting energy expenditure are small, insignificant, or
not physiologically relevant.
\end{abstract}

@@

Studies were carried out in multiple countries: Brazil (k = 3) ,and USA
(k = 3 studies), Australia, Cameroon, Canada, Italy, Sweden, Turkey, UK
(all k = 1 study). The four studies with noted reporting errors were
carried out in Greece. A total of 11 studies used the Rotterdam criteria
(or a modified version there of) for diagnosing
PCOS\textsuperscript{16,17,34,53--60}. One study used the 1990 National
Institutes of Health criteria\textsuperscript{61}, and two
studies\textsuperscript{18,55} diagnosed PCOS via the presence of
oligomenorrhea or amenorrheaammenorrhea alongside additional criteria
including plasma androgen levels, hirsutism or polycystic ovaries on
ultrasound scanning. All but one study\textsuperscript{61}, which used
doubly labelled water, measured REE using indirect calorimetry. The
specific devices reported by these studies are included in the online
supplementary materials (see
\href{https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html}{Descriptives
Table}). We originally considered in our pre-registration that, given
sufficient data, we would compare sub groups of women with PCOS who did
and did not have accompanying insulin resistance. However, based on the
metabolic health variables reported in studies with this information
(see
\href{https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html}{Descriptives
Table}) and, where available, considering primary criteria of either
homeostatic model assessment of insulin resistance (HOMA-IR) \(\geq2.5\)
or secondary criteria including fasting insulin \(>12\mu \textrm{U/mL}\)
or fasting glucose \(\geq100\textrm{mg/dL}\), all groups of women with
PCOS in the included studies would be considered to have insulin
resistance. Mean age of women with PCOS in these studies ranged from 23
to 33 which was similar to the control women without PCOS ranging from
23 to 30. Across those studies where BMI was reported or it was possible
to estimate the mean BMI of women with PCOS ranged from 26.4 to 39.9 was
typically greater than women without PCOS which ranged 20.5 to 27.9.

@@

These findings challenge the widely held belief that PCOS is inherently
associated with a slower metabolism (PMID: 32244780), predisposing women
with PCOS to weight gain. This belief may mistakenly lead to
recommendations centred on slightly more severe calorie restriction to
achieve weight-loss goals, compared with recommendations for the general
population, as primary management strategies for women with
PCOS\textsuperscript{20}. Recognising that there may be minimal
differences in REE between women with PCOS and those without can inform
both clinical and public practices, potentially leading to a shift in
focus away from a requirement for more severe caloric restriction as a
primary method of treatment towards more comprehensive, individualised,
and psychologically safe approaches to care\textsuperscript{25}.

The current study suggests that REE may not a barrier to weight
regulation in PCOS given small group-level differences between women
with PCOS and those without (229 kcal/day {[}95\%QI:169 kcal/day to 303
kcal/day{]} versus 238 kcal/day {[}95\%QI:178 kcal/day to 312
kcal/day{]}, respectively). If anything, REE may be slightly higher in
women with PCOS compared to healthy women without PCOS. BMI of women
with PCOS ranged from 26.4 to 39.9 was typically greater than women
without PCOS which ranged 20.5 to 27.9 and this may explain the slightly
greater REE in the former group. However, for those studies where we
could extract or estimate BMI, our additional exploratory models
adjusted for this similarly showed little difference in REE (20 kcal/day
{[}95\%QI: -96 kcal/day to 140kcal/day{]}) between women with PCOS and
those without. However, one of the studies included in our
analysis\textsuperscript{62} reported that, whilst there was little
difference in unadjusted REE, when REE was adjusted for fat-free mass it
was lower in women with PCOS suggesting the potential importance of
fat-free tissue in energy regulation. Yet, for those studies where
fat-free mass was reported, our additional exploratory models adjusted
for this similarly showed little difference in REE (-9 kcal/day
{[}95\%QI: -195 kcal/day to 168kcal/day{]}) between women with PCOS and
those without. As such, even adjusted for both BMI and fat-free mass,
there seems to be little difference in REE between women with, and
without, PCOS. Yet, a recent systematic review of mechanisms for
metabolic dysfunction has reported excess androgen drives metabolic
issues within adipose tissue and muscle tissue contributing to
complications like obesity and insulin resistance (PMID: 32244780).
Taken together, these findings highlight that factors beyond REE and
typical correlates of this including BMI or fat-free mass, such as
hormonal and tissue-specific metabolic effects, may play a more
significant role in weight regulation challenges in women with PCOS.

As noted, women with PCOS are more likely to engage in weight-loss
attempts\textsuperscript{19} and there could be concerns that this could
inadvertently further foster the already well documented disordered
eating in this population\textsuperscript{21,22,63,64}. The pathways
linking PCOS and disordered eating are multifactorial. Biological
mechanisms such as hyperandrogenism, hyperinsulinaemia, and altered
ghrelin and leptin signalling can heighten hunger, carbohydrate
cravings, and appetite variability\textsuperscript{65}. Frequent
hypoglycaemia and associated mood changes have also been observed, which
can trigger compensatory eating or binge episodes\textsuperscript{65}.
These physiological processes interact with psychological and social
stressors, including infertility concerns, conflicting nutrition advice,
chronic dieting, and exposure to idealised body images on social media,
which together compound vulnerability to disordered
eating\textsuperscript{66}. Moreover, eating disorders themselves can
disrupt endocrine function, potentially worsening PCOS symptoms and
creating a self-reinforcing cycle\textsuperscript{64,67}.

Understanding the intertwined biological, psychological, and social
influences suggests the importance of considering whether restrictive
dietary advice is appropriate given it may exacerbate feelings of
failure, hunger dysregulation, and shame\textsuperscript{68}. These
concerns, coupled with the lack of difference in REE between women with
and without PCOS might suggest that energy restriction based dietary
interventions for weight-loss may be unnecessary. But, there is also
evidence supporting the effect of energy restricted dietary
interventions for improving PCOS symptoms\textsuperscript{23,24} and
they are recommended in international guidelines\textsuperscript{25}.
Encouragingly though, these guidelines also recognise weight stigma as a
determinant of health and call for its reduction across clinical and
public health settings. Evidently the greater prevalence of women with
PCOS falling into overweight and obese BMI categories compared to women
without PCOS\textsuperscript{8,9} is unlikely to be due to differences
in REE and so, despite the potential effectiveness of energy restriction
dietary interventions, there is potential value in moving towards more
weight-neutral, individualised, and empowering care following holistic
guidelines recommending multiple approaches to
management\textsuperscript{20,25}.

@@

Hannah Cabré was supported by the National Institute of Diabetes and
Digestive and Kidney Diseases of the National Institutes of Health under
Award Number T32DK064584. James Steele provides research consultancy
through his company Steele Research Limited, is contracted currently by
MacroFactor and Kieser Australia through Steele Research Limited, and
has also received travel expenses and honoraria for speaking from fit20
International, Exercise School Portugal, and Discover Strength.

The content is solely the responsibility of the authors and does not
necessarily represent the official views of the National Institutes of
Health or any other funding agencies or institutions noted above.

@@

\bibitem[\citeproctext]{ref-leeIncreasedRiskDisordered2017}
\CSLLeftMargin{63. }%
\CSLRightInline{Lee I, Cooney LG, Saini S, et al. Increased risk of
disordered eating in~polycystic ovary syndrome. \emph{Fertility and
Sterility}. 2017;107(3):796-802.
doi:\href{https://doi.org/10.1016/j.fertnstert.2016.12.014}{10.1016/j.fertnstert.2016.12.014}}

\bibitem[\citeproctext]{ref-bernadettPrevalenceEatingDisorders2016}
\CSLLeftMargin{64. }%
\CSLRightInline{Bernadett M, Szemán-N A.
\href{https://www.ncbi.nlm.nih.gov/pubmed/27244869}{{{[}Prevalence of
eating disorders among women with polycystic ovary syndrome{]}}}.
\emph{Psychiatria Hungarica: A Magyar Pszichiatriai Tarsasag Tudomanyos
Folyoirata}. 2016;31(2):136-145.}

\bibitem[\citeproctext]{ref-barryImpactEatingBehavior2011}
\CSLLeftMargin{65. }%
\CSLRightInline{Barry JA, Bouloux P, Hardiman PJ. The impact of eating
behavior on psychological symptoms typical of reactive hypoglycemia. {A}
pilot study comparing women with polycystic ovary syndrome to controls.
\emph{Appetite}. 2011;57(1):73-76.
doi:\href{https://doi.org/10.1016/j.appet.2011.03.003}{10.1016/j.appet.2011.03.003}}

\bibitem[\citeproctext]{ref-steegers-theunissenPolycysticOvarySyndrome2020}
\CSLLeftMargin{66. }%
\CSLRightInline{Steegers-Theunissen RPM, Wiegel RE, Jansen PW, Laven
JSE, Sinclair KD. Polycystic {Ovary Syndrome}: {A Brain Disorder
Characterized} by {Eating Problems Originating} during {Puberty} and
{Adolescence}. \emph{International Journal of Molecular Sciences}.
2020;21(21):8211.
doi:\href{https://doi.org/10.3390/ijms21218211}{10.3390/ijms21218211}}

\bibitem[\citeproctext]{ref-cooneyIncreasedPrevalenceBinge2024}
\CSLLeftMargin{67. }%
\CSLRightInline{Cooney LG, Gyorfi K, Sanneh A, et al. Increased
{Prevalence} of {Binge Eating Disorder} and {Bulimia Nervosa} in {Women
With Polycystic Ovary Syndrome}: {A Systematic Review} and
{Meta-Analysis}. \emph{The Journal of Clinical Endocrinology and
Metabolism}. 2024;109(12):3293-3305.
doi:\href{https://doi.org/10.1210/clinem/dgae462}{10.1210/clinem/dgae462}}

\bibitem[\citeproctext]{ref-pirottaDisorderedEatingBehaviours2019}
\CSLLeftMargin{68. }%
\CSLRightInline{Pirotta S, Barillaro M, Brennan L, et al. Disordered
{Eating Behaviours} and {Eating Disorders} in {Women} in {Australia}
with and without {Polycystic Ovary Syndrome}: {A Cross-Sectional Study}.
\emph{Journal of Clinical Medicine}. 2019;8(10):1682.
doi:\href{https://doi.org/10.3390/jcm8101682}{10.3390/jcm8101682}}

>>>>

\usepackage[font=scriptsize]{caption}
\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\usepackage{etoolbox}
\AtEndEnvironment{abstract}{\noindent\textbf{Keywords:} \KeywordList \par}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Resting energy expenditure of women with and without polycystic ovary syndrome: a systematic review and meta-analysis},
  pdfauthor={Richard Kirwan; Leigh Peele; Gregory Nuckols; Georgia Kohlhoff; Hannah Cabré; Alyssa Olenick; James Steele},
  pdfkeywords={Polycystic ovary syndrome, Energy Metabolism, Resting
Metabolic Rate, Basal Metabolism},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Resting energy expenditure of women with and without polycystic
ovary syndrome: a systematic review and meta-analysis\thanks{Preprint,
please cite as: Kirwan, R., Peele, L., Nuckols, G., Kohlhoff, G., Cabré,
H., Olenick, A., and Steele, J. (2025). Resting energy expenditure of
women with and without polycystic ovary syndrome: a systematic review
and meta-analysis. medrxiv DOI: TO ADD. Address for correspondence:
james@steele-research.com}}

\author[1]{Richard Kirwan}
\author[2]{Leigh Peele}
\author[2]{Gregory Nuckols}
\author[3]{Georgia Kohlhoff}
\author[4]{Hannah Cabré}
\author[5]{Alyssa Olenick}
\author[2,6,7]{James Steele}

\affil[1]{Research Institute for Sports and Exercise Sciences, Liverpool
John Moores University, Liverpool, UK}
\affil[2]{MacroFactor, Stronger by Science Technologies LLC, Raleigh,
North Carolina, USA}
\affil[3]{Flourishing Health, Liverpool, UK}
\affil[4]{Reproductive Endocrinology and Women's Health Research
Program, Pennington Biomedical Research Center, Baton Rouge, Louisiana,
USA}
\affil[5]{Dr Alyssa Olenick LLC}
\affil[6]{Steele Research Limited, Eastleigh, Hampshire, UK}
\affil[7]{School of Health and Biomedical Sciences, Royal Melbourne
Institute of Technology, Melbourne, VIC, Australia}

% Store the keywords (expanded by Pandoc here)
\def\KeywordList{Polycystic ovary syndrome, Energy Metabolism, Resting
Metabolic Rate, Basal Metabolism}

\date{December 3, 2025}
\begin{document}
\maketitle
\begin{abstract}
Context: Polycystic ovary syndrome (PCOS) is common in reproductive-age
women, who often have higher BMI classification. This is assumed to stem
from lower resting energy expenditure (REE), influencing lifestyle
intervention guidelines. However, evidence for reduced REE in women with
PCOS compared with those without is inconsistent. Objective: To
systematically search and meta-analyse the existing literature to
estimate and describe the difference in REE between women with and
without PCOS. Data Sources: A systematic search was conducted using
PubMed, Medline and Web of Science databases of published research from
January 1990 to January 2025. Study Selection: Studies that measured REE
in women living with PCOS, both with and without control arms of women
without PCOS, were included. Data Extraction: Bibliometric, demographic,
and REE data was extracted by one investigator and checked in
triplicate. Data Synthesis: Thirteen studies were included in a Bayesian
arm-based multiple condition comparison (i.e., network) type
meta-analysis model with informative priors to compare both mean REE,
and between person variation in REE, between women with and without
PCOS. Mean REE differed between groups by 30 kcal/day {[}95\% quantile
interval: -47 to 113 kcal/day{]} and the contrast ratio for between
person standard deviations was 0.98 {[}95\% quantile interval: 0.71 to
1.33{]}. Conclusions: These findings indicate that REE does not
meaningfully differ between women with and without PCOS. Group-level
differences in resting energy expenditure are small, insignificant, or
not physiologically relevant.
\end{abstract}

@@

Studies were carried out in multiple countries: Brazil (k = 3) ,and USA
(k = 3 studies), Australia, Cameroon, Canada, Italy, Sweden, Turkey, UK
(all k = 1 study). The four studies with noted reporting errors were
carried out in Greece. A total of 11 studies used the Rotterdam criteria
(or a modified version there of) for diagnosing
PCOS\textsuperscript{16,17,34,53--60}. One study used the 1990 National
Institutes of Health criteria\textsuperscript{61}, and two
studies\textsuperscript{18,55} diagnosed PCOS via the presence of
oligomenorrhea or amenorrhea alongside additional criteria including
plasma androgen levels, hirsutism or polycystic ovaries on ultrasound
scanning. All but one study\textsuperscript{61}, which used doubly
labelled water, measured REE using indirect calorimetry. The specific
devices reported by these studies are included in the online
supplementary materials (see
\href{https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html}{Descriptives
Table}). We originally considered in our pre-registration that, given
sufficient data, we would compare sub groups of women with PCOS who did
and did not have accompanying insulin resistance. However, based on the
metabolic health variables reported in studies with this information
(see
\href{https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html}{Descriptives
Table}) and, where available, considering primary criteria of either
homeostatic model assessment of insulin resistance (HOMA-IR) \(\geq2.5\)
or secondary criteria including fasting insulin \(>12\mu \textrm{U/mL}\)
or fasting glucose \(\geq100\textrm{mg/dL}\), all groups of women with
PCOS in the included studies would be considered to have insulin
resistance. Mean age of women with PCOS in these studies ranged from 23
to 33 which was similar to the control women without PCOS ranging from
23 to 30. Across those studies where BMI was reported or it was possible
to estimate the mean BMI of women with PCOS ranged from 26.4 to 39.9 was
typically greater than women without PCOS which ranged 20.5 to 27.9.

@@

These findings challenge the widely held belief that PCOS is inherently
associated with a slower metabolism\textsuperscript{@ 25}, predisposing
women with PCOS to weight gain. This belief largely stems from a single
but influential 2009 study from Georgopoulos et al.~that reported a
significantly reduced BMR in women with PCOS\textsuperscript{13}, which
has been widely cited and reinforced in both academic and clinical
contexts. However, as we have noted, this study along with
others\textsuperscript{35--37} have numerous reporting errors which led
us to drop them from our present analysis (though sensitivity analysis
including them did not alter our overall conclusions). The mistaken
belief that REE is lower for those with PCOS may have mistakenly lead to
recommendations centred on slightly more severe calorie restriction to
achieve weight-loss goals, compared with recommendations for the general
population, as primary management strategies for women with
PCOS\textsuperscript{20}. Recognising that there may be minimal
differences in REE between women with PCOS and those without can inform
both clinical and public practices, potentially leading to a shift in
focus away from a requirement for more severe caloric restriction as a
primary method of treatment towards more comprehensive, individualised,
and psychologically safe approaches to care\textsuperscript{25}.

The current study suggests that REE may not a barrier to weight
regulation in PCOS given small group-level differences between women
with PCOS and those without (1472 kcal/day {[}95\%QI:1359 kcal/day to
1587 kcal/day{]} versus 1442 kcal/day {[}95\%QI:1334 kcal/day to 1553
kcal/day{]}, respectively). If anything, REE may be slightly higher in
women with PCOS compared to healthy women without PCOS. BMI of women
with PCOS ranged from 26.4 to 39.9 was typically greater than women
without PCOS which ranged 20.5 to 27.9 and this may explain the slightly
greater REE in the former group. However, for those studies where we
could extract or estimate BMI, our additional exploratory models
adjusted for this similarly showed little difference in REE (20 kcal/day
{[}95\%QI: -96 kcal/day to 140kcal/day{]}) between women with PCOS and
those without. However, one of the studies included in our
analysis\textsuperscript{62} reported that, whilst there was little
difference in unadjusted REE, when REE was adjusted for fat-free mass it
was lower in women with PCOS suggesting the potential importance of
fat-free tissue in energy regulation. Yet, for those studies where
fat-free mass was reported, our additional exploratory models adjusted
for this similarly showed little difference in REE (-9 kcal/day
{[}95\%QI: -195 kcal/day to 168kcal/day{]}) between women with PCOS and
those without. As such, even adjusted for both BMI and fat-free mass,
there seems to be little difference in REE between women with, and
without, PCOS. Yet, a recent systematic review of mechanisms for
metabolic dysfunction has reported excess androgen drives metabolic
issues within adipose tissue and muscle tissue contributing to
complications like obesity and insulin resistance\textsuperscript{63}.
Taken together, these findings highlight that factors beyond REE and
typical correlates of this including BMI or fat-free mass, such as
hormonal and tissue-specific metabolic effects, may play a more
significant role in weight regulation challenges in women with PCOS.

As noted, women with PCOS are more likely to engage in weight-loss
attempts\textsuperscript{19} and there could be concerns that this could
inadvertently further foster the already well documented disordered
eating in this population\textsuperscript{21,22,64,65}. The pathways
linking PCOS and disordered eating are multifactorial. Biological
mechanisms such as hyperandrogenism, hyperinsulinaemia, and altered
ghrelin and leptin signalling can heighten hunger, carbohydrate
cravings, and appetite variability\textsuperscript{66}. Frequent
hypoglycaemia and associated mood changes have also been observed, which
can trigger compensatory eating or binge episodes\textsuperscript{66}.
These physiological processes interact with psychological and social
stressors, including infertility concerns, conflicting nutrition advice,
chronic dieting, and exposure to idealised body images on social media,
which together compound vulnerability to disordered
eating\textsuperscript{67}. Moreover, eating disorders themselves can
disrupt endocrine function, potentially worsening PCOS symptoms and
creating a self-reinforcing cycle\textsuperscript{65,68}.

Understanding the intertwined biological, psychological, and social
influences suggests the importance of considering whether restrictive
dietary advice is appropriate given it may exacerbate feelings of
failure, hunger dysregulation, and shame\textsuperscript{69}. These
concerns, coupled with the lack of difference in REE between women with
and without PCOS might suggest that energy restriction based dietary
interventions for weight-loss may be unnecessary. But, there is also
evidence supporting the effect of energy restricted dietary
interventions for improving PCOS symptoms\textsuperscript{23,24} and
they are recommended in international guidelines\textsuperscript{25}.
Encouragingly though, these guidelines also recognise weight stigma as a
determinant of health and call for its reduction across clinical and
public health settings. Evidently the greater prevalence of women with
PCOS falling into overweight and obese BMI categories compared to women
without PCOS\textsuperscript{8,9} is unlikely to be due to differences
in REE and so, despite the potential effectiveness of energy restriction
dietary interventions, there is potential value in moving towards more
weight-neutral, individualised, and empowering care following holistic
guidelines recommending multiple approaches to
management\textsuperscript{20,25}.

@@

Richard Kirwan provides nutrition consultancy through his company Be
More Nutrition. Leigh Peele is employed by MacroFactor, a
nutrition-tracking app. Gregory Nuckols is co-owner of MacroFactor.
Georgia Kohlhoff provides nutrition consultancy through her company
Flourishing Health. Hannah Cabré was supported by the National Institute
of Diabetes and Digestive and Kidney Diseases of the National Institutes
of Health under Award Number T32DK064584. Alyssa Olenick provides
fitness coaching and scientific consulting under ALYSSA OLENICK LLC. She
receives compensation from scientific consulting, speaking engagments,
affiliated social media partnerships, and book publishing. James Steele
provides research consultancy through his company Steele Research
Limited, is contracted currently by MacroFactor and Kieser Australia
through Steele Research Limited, and has also received travel expenses
and honoraria for speaking from fit20 International, Exercise School
Portugal, and Discover Strength.

The content is solely the responsibility of the authors and does not
necessarily represent the official views of the National Institutes of
Health or any other funding agencies or institutions noted above. The
authors declare no other conflicts of interest related to the submitted
work.

@@

\bibitem[\citeproctext]{ref-sanchez-garridoMetabolicDysfunctionPolycystic2020}
\CSLLeftMargin{63. }%
\CSLRightInline{Sanchez-Garrido MA, Tena-Sempere M. Metabolic
dysfunction in polycystic ovary syndrome: {Pathogenic} role of androgen
excess and potential therapeutic strategies. \emph{Molecular
Metabolism}. 2020;35:100937.
doi:\href{https://doi.org/10.1016/j.molmet.2020.01.001}{10.1016/j.molmet.2020.01.001}}

\bibitem[\citeproctext]{ref-leeIncreasedRiskDisordered2017}
\CSLLeftMargin{64. }%
\CSLRightInline{Lee I, Cooney LG, Saini S, et al. Increased risk of
disordered eating in~polycystic ovary syndrome. \emph{Fertility and
Sterility}. 2017;107(3):796-802.
doi:\href{https://doi.org/10.1016/j.fertnstert.2016.12.014}{10.1016/j.fertnstert.2016.12.014}}

\bibitem[\citeproctext]{ref-bernadettPrevalenceEatingDisorders2016}
\CSLLeftMargin{65. }%
\CSLRightInline{Bernadett M, Szemán-N A.
\href{https://www.ncbi.nlm.nih.gov/pubmed/27244869}{{{[}Prevalence of
eating disorders among women with polycystic ovary syndrome{]}}}.
\emph{Psychiatria Hungarica: A Magyar Pszichiatriai Tarsasag Tudomanyos
Folyoirata}. 2016;31(2):136-145.}

\bibitem[\citeproctext]{ref-barryImpactEatingBehavior2011}
\CSLLeftMargin{66. }%
\CSLRightInline{Barry JA, Bouloux P, Hardiman PJ. The impact of eating
behavior on psychological symptoms typical of reactive hypoglycemia. {A}
pilot study comparing women with polycystic ovary syndrome to controls.
\emph{Appetite}. 2011;57(1):73-76.
doi:\href{https://doi.org/10.1016/j.appet.2011.03.003}{10.1016/j.appet.2011.03.003}}

\bibitem[\citeproctext]{ref-steegers-theunissenPolycysticOvarySyndrome2020}
\CSLLeftMargin{67. }%
\CSLRightInline{Steegers-Theunissen RPM, Wiegel RE, Jansen PW, Laven
JSE, Sinclair KD. Polycystic {Ovary Syndrome}: {A Brain Disorder
Characterized} by {Eating Problems Originating} during {Puberty} and
{Adolescence}. \emph{International Journal of Molecular Sciences}.
2020;21(21):8211.
doi:\href{https://doi.org/10.3390/ijms21218211}{10.3390/ijms21218211}}

\bibitem[\citeproctext]{ref-cooneyIncreasedPrevalenceBinge2024}
\CSLLeftMargin{68. }%
\CSLRightInline{Cooney LG, Gyorfi K, Sanneh A, et al. Increased
{Prevalence} of {Binge Eating Disorder} and {Bulimia Nervosa} in {Women
With Polycystic Ovary Syndrome}: {A Systematic Review} and
{Meta-Analysis}. \emph{The Journal of Clinical Endocrinology and
Metabolism}. 2024;109(12):3293-3305.
doi:\href{https://doi.org/10.1210/clinem/dgae462}{10.1210/clinem/dgae462}}

\bibitem[\citeproctext]{ref-pirottaDisorderedEatingBehaviours2019}
\CSLLeftMargin{69. }%
\CSLRightInline{Pirotta S, Barillaro M, Brennan L, et al. Disordered
{Eating Behaviours} and {Eating Disorders} in {Women} in {Australia}
with and without {Polycystic Ovary Syndrome}: {A Cross-Sectional Study}.
\emph{Journal of Clinical Medicine}. 2019;8(10):1682.
doi:\href{https://doi.org/10.3390/jcm8101682}{10.3390/jcm8101682}}
